---
title: "Figure 3"
author: "Lu Zeng & Steve Pederson"
date: "8 March 2016"
output: html_document
---

# Introduction
This script analyses and prepares the data for Figure 3.

In this analysis, the analysed values are the probability of a nucleotide being within an RE, given that it is within an TE, i.e. $p(RE \mid TE)$.
The TEs are explored using the specific classes:

1. SINE/Alu
2. SINE/Mir
3. LINE/L1 Ancestral
4. LINE/L1 Recent
5. LINE/L2
6. LTR/ERV1         
7. DNA/hAT-Charlie 

# Data Preparation

```{r, message=FALSE, warning=FALSE}
library(ggplot2)
library(MASS)
library(matrixStats)
library(xtable)
library(multcomp)
library(sandwich)
library(car)
library(boot)
library(dplyr)
library(reshape2)
library(readr)
library(knitr)
library(magrittr)
library(limma)
library(spShortcuts)
library(grid)
```

```{r}
fig3 <- file.path("data", "Figure3.csv") %>%
  read_csv(col_names = FALSE) %>%
  set_colnames(c("TE", "RE", "Tissue", "BP")) %>%
  mutate(TE = factor(TE, unique(TE)),
         RE = capwords(RE),
         RE = gsub("Regions", "Region", RE),
         RE = factor(RE, unique(RE)),
         Tissue = capwords(Tissue),
         BP = BP / 100,
         Class = gsub("(.+)/.+", "\\1", TE),
         Type = gsub(".+/(.+)", "\\1", TE),
         shortTE = gsub("(L|S)INE/", "", TE),
         shortTE = gsub("(LTR|DNA).+", "\\1", shortTE),
         shortTE = factor(shortTE, unique(shortTE)),
         logitBP = logit(BP/100))
```

# Raw Data

```{r, echo=FALSE, fig.height=5, fig.width = 8, fig.cap = 'Conditional probabilities for a nucleotide being an RE, given that it is within a TE'}
fs <- 8
fig3a <- fig3 %>%
  group_by(RE, TE) %>%
  summarise(mn = mean(logitBP), n = n(), se = sd(logitBP) /sqrt(n)) %>%
  ungroup()%>%
  mutate(lwr = binomial()$linkinv(mn - se),
         upr = binomial()$linkinv(mn + se),
         pi = binomial()$linkinv(mn),
         RE  = gsub(" ", "\n", RE),
         RE = factor(RE, unique(RE))) %>%
  ggplot(aes(x = TE, y = pi, fill = TE)) +
  geom_bar(stat = "identity") +
  geom_errorbar(aes(ymin = lwr, ymax = upr), width=  0.4) +
  facet_wrap(~RE, nrow = 1) +
  theme_bw() +
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(x = c(),
       y = "Probability Estimate",
       fill = "Element\nType") +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.y = element_text(size = fs),
        axis.text.y = element_text(size = fs-1),
        strip.text = element_text(size = fs-1),
        legend.text = element_text(size = fs),
        legend.title = element_text(size = fs, face = "bold"),
        legend.title.align = 0.5,
        legend.position = "top",
        panel.margin = unit(.004, "npc"),
        panel.grid = element_line(colour = "grey50")) 
fig3a
```


# Model Fitting and Testing

The model was fitting as a whole dataset and whilst the assumption of normality did hold, whilst the assumption of homoscedasticity did not.
```{r, fig3Lm}
fig3Lm <- lm(logitBP ~ shortTE*RE, data = fig3)
```

```{r, swTest}
fig3Lm %>% 
  resid() %>%
  shapiro.test()
```

```{r, levTest}
fig3Lm %>%
  leveneTest(center = "mean")
```

```{r, echo=FALSE, include=TRUE}
fig3Lm %>%
  glht(vcov= sandwich) %>%
  summary(test = adjusted("bonferroni")) %>%
  extract2("test") %>%
  extract(c("coefficients", "sigma", "tstat", "pvalues")) %>%
  as.data.frame() %>%
  add_rownames("Coefficient") %>%
  rename(Estimate = coefficients,
         `Std Error` = sigma,
         T = tstat,
         P = pvalues) %>%
  mutate(Coefficient = gsub("(shortTE|RE)", "", Coefficient),
         ` ` = aster(P),
         P = format(P, digits= 3, scientific = TRUE)) %>%
  kable(digits = 3,
        caption = "Coefficients from model fitting. P values are adjusted using Bonferroni's Method",
        align = c("l", "r", "r", "r", "r", "l"))
```


# Preparation of Contrasts

A matrix of all possible contrasts for each RE was then formed based on the full model
```{r, contMat}
modMat <- fig3Lm %>%
  update(.~. -1) %>%
  model.matrix
colnames(modMat) %<>% 
  magsub("(shortTE|RE)", "") %>%
  magsub(" ", "_") %>%
  magsub(":", ".")
```

```{r}
contMat <- list(
  `Active Promoter` = makeContrasts(
    Mir - Alu,
    L1_Ancestral - Alu,
    L1_Recent - Alu,
    L2 - Alu,
    LTR - Alu,
    DNA - Alu,
    L1_Ancestral - Mir,
    L1_Recent - Mir,
    L2 - Mir,
    LTR - Mir,
    DNA - Mir,
    L1_Recent - L1_Ancestral,
    L2 - L1_Ancestral,
    LTR - L1_Ancestral,
    DNA - L1_Ancestral,
    L2 - L1_Recent,
    LTR - L1_Recent,
    DNA - L1_Recent,
    LTR - L2,
    DNA - L2,
    DNA - LTR,
    levels = modMat),
  `Weak Promoter` = makeContrasts(
    Mir + Mir.Weak_Promoter - Alu,
    L1_Ancestral + L1_Ancestral.Weak_Promoter - Alu,
    L1_Recent + L1_Recent.Weak_Promoter - Alu,
    L2 + L2.Weak_Promoter - Alu,
    LTR + LTR.Weak_Promoter - Alu,
    DNA + DNA.Weak_Promoter - Alu,
    L1_Ancestral + L1_Ancestral.Weak_Promoter - Mir - Mir.Weak_Promoter,
    L1_Recent + L1_Recent.Weak_Promoter - Mir - Mir.Weak_Promoter,
    L2 + L2.Weak_Promoter- Mir - Mir.Weak_Promoter,
    LTR + LTR.Weak_Promoter - Mir - Mir.Weak_Promoter,
    DNA + DNA.Weak_Promoter - Mir - Mir.Weak_Promoter,
    L1_Recent + L1_Recent.Weak_Promoter - L1_Ancestral - L1_Ancestral.Weak_Promoter,
    L2 + L2.Weak_Promoter - L1_Ancestral - L1_Ancestral.Weak_Promoter,
    LTR + LTR.Weak_Promoter - L1_Ancestral - L1_Ancestral.Weak_Promoter,
    DNA + DNA.Weak_Promoter - L1_Ancestral - L1_Ancestral.Weak_Promoter,
    L2 + L2.Weak_Promoter - L1_Recent - L1_Recent.Weak_Promoter,
    LTR + LTR.Weak_Promoter - L1_Recent - L1_Recent.Weak_Promoter,
    DNA + DNA.Weak_Promoter - L1_Recent - L1_Recent.Weak_Promoter,
    LTR + LTR.Weak_Promoter - L2 - L2.Weak_Promoter,
    DNA + DNA.Weak_Promoter - L2 - L2.Weak_Promoter,
    DNA + DNA.Weak_Promoter - LTR - LTR.Weak_Promoter,
    levels = modMat),
  `Strong Enhancer` = makeContrasts(
    Mir + Mir.Strong_Enhancer - Alu,
    L1_Ancestral + L1_Ancestral.Strong_Enhancer - Alu,
    L1_Recent + L1_Recent.Strong_Enhancer - Alu,
    L2 + L2.Strong_Enhancer - Alu,
    LTR + LTR.Strong_Enhancer - Alu,
    DNA + DNA.Strong_Enhancer - Alu,
    L1_Ancestral + L1_Ancestral.Strong_Enhancer - Mir - Mir.Strong_Enhancer,
    L1_Recent + L1_Recent.Strong_Enhancer - Mir - Mir.Strong_Enhancer,
    L2 + L2.Strong_Enhancer- Mir - Mir.Strong_Enhancer,
    LTR + LTR.Strong_Enhancer - Mir - Mir.Strong_Enhancer,
    DNA + DNA.Strong_Enhancer - Mir - Mir.Strong_Enhancer,
    L1_Recent + L1_Recent.Strong_Enhancer - L1_Ancestral - L1_Ancestral.Strong_Enhancer,
    L2 + L2.Strong_Enhancer - L1_Ancestral - L1_Ancestral.Strong_Enhancer,
    LTR + LTR.Strong_Enhancer - L1_Ancestral - L1_Ancestral.Strong_Enhancer,
    DNA + DNA.Strong_Enhancer - L1_Ancestral - L1_Ancestral.Strong_Enhancer,
    L2 + L2.Strong_Enhancer - L1_Recent - L1_Recent.Strong_Enhancer,
    LTR + LTR.Strong_Enhancer - L1_Recent - L1_Recent.Strong_Enhancer,
    DNA + DNA.Strong_Enhancer - L1_Recent - L1_Recent.Strong_Enhancer,
    LTR + LTR.Strong_Enhancer - L2 - L2.Strong_Enhancer,
    DNA + DNA.Strong_Enhancer - L2 - L2.Strong_Enhancer,
    DNA + DNA.Strong_Enhancer - LTR - LTR.Strong_Enhancer,
    levels = modMat),
  `Weak Enhancer` = makeContrasts(
    Mir + Mir.Weak_Enhancer - Alu,
    L1_Ancestral + L1_Ancestral.Weak_Enhancer - Alu,
    L1_Recent + L1_Recent.Weak_Enhancer - Alu,
    L2 + L2.Weak_Enhancer - Alu,
    LTR + LTR.Weak_Enhancer - Alu,
    DNA + DNA.Weak_Enhancer - Alu,
    L1_Ancestral + L1_Ancestral.Weak_Enhancer - Mir - Mir.Weak_Enhancer,
    L1_Recent + L1_Recent.Weak_Enhancer - Mir - Mir.Weak_Enhancer,
    L2 + L2.Weak_Enhancer- Mir - Mir.Weak_Enhancer,
    LTR + LTR.Weak_Enhancer - Mir - Mir.Weak_Enhancer,
    DNA + DNA.Weak_Enhancer - Mir - Mir.Weak_Enhancer,
    L1_Recent + L1_Recent.Weak_Enhancer - L1_Ancestral - L1_Ancestral.Weak_Enhancer,
    L2 + L2.Weak_Enhancer - L1_Ancestral - L1_Ancestral.Weak_Enhancer,
    LTR + LTR.Weak_Enhancer - L1_Ancestral - L1_Ancestral.Weak_Enhancer,
    DNA + DNA.Weak_Enhancer - L1_Ancestral - L1_Ancestral.Weak_Enhancer,
    L2 + L2.Weak_Enhancer - L1_Recent - L1_Recent.Weak_Enhancer,
    LTR + LTR.Weak_Enhancer - L1_Recent - L1_Recent.Weak_Enhancer,
    DNA + DNA.Weak_Enhancer - L1_Recent - L1_Recent.Weak_Enhancer,
    LTR + LTR.Weak_Enhancer - L2 - L2.Weak_Enhancer,
    DNA + DNA.Weak_Enhancer - L2 - L2.Weak_Enhancer,
    DNA + DNA.Weak_Enhancer - LTR - LTR.Weak_Enhancer,
    levels = modMat),
  `Repressed Region` = makeContrasts(
    Mir + Mir.Repressed_Region - Alu,
    L1_Ancestral + L1_Ancestral.Repressed_Region - Alu,
    L1_Recent + L1_Recent.Repressed_Region - Alu,
    L2 + L2.Repressed_Region - Alu,
    LTR + LTR.Repressed_Region - Alu,
    DNA + DNA.Repressed_Region - Alu,
    L1_Ancestral + L1_Ancestral.Repressed_Region - Mir - Mir.Repressed_Region,
    L1_Recent + L1_Recent.Repressed_Region - Mir - Mir.Repressed_Region,
    L2 + L2.Repressed_Region- Mir - Mir.Repressed_Region,
    LTR + LTR.Repressed_Region - Mir - Mir.Repressed_Region,
    DNA + DNA.Repressed_Region - Mir - Mir.Repressed_Region,
    L1_Recent + L1_Recent.Repressed_Region - L1_Ancestral - L1_Ancestral.Repressed_Region,
    L2 + L2.Repressed_Region - L1_Ancestral - L1_Ancestral.Repressed_Region,
    LTR + LTR.Repressed_Region - L1_Ancestral - L1_Ancestral.Repressed_Region,
    DNA + DNA.Repressed_Region - L1_Ancestral - L1_Ancestral.Repressed_Region,
    L2 + L2.Repressed_Region - L1_Recent - L1_Recent.Repressed_Region,
    LTR + LTR.Repressed_Region - L1_Recent - L1_Recent.Repressed_Region,
    DNA + DNA.Repressed_Region - L1_Recent - L1_Recent.Repressed_Region,
    LTR + LTR.Repressed_Region - L2 - L2.Repressed_Region,
    DNA + DNA.Repressed_Region - L2 - L2.Repressed_Region,
    DNA + DNA.Repressed_Region - LTR - LTR.Repressed_Region,
    levels = modMat),
  `Insulator` = makeContrasts(
    Mir + Mir.Insulator - Alu,
    L1_Ancestral + L1_Ancestral.Insulator - Alu,
    L1_Recent + L1_Recent.Insulator - Alu,
    L2 + L2.Insulator - Alu,
    LTR + LTR.Insulator - Alu,
    DNA + DNA.Insulator - Alu,
    L1_Ancestral + L1_Ancestral.Insulator - Mir - Mir.Insulator,
    L1_Recent + L1_Recent.Insulator - Mir - Mir.Insulator,
    L2 + L2.Insulator- Mir - Mir.Insulator,
    LTR + LTR.Insulator - Mir - Mir.Insulator,
    DNA + DNA.Insulator - Mir - Mir.Insulator,
    L1_Recent + L1_Recent.Insulator - L1_Ancestral - L1_Ancestral.Insulator,
    L2 + L2.Insulator - L1_Ancestral - L1_Ancestral.Insulator,
    LTR + LTR.Insulator - L1_Ancestral - L1_Ancestral.Insulator,
    DNA + DNA.Insulator - L1_Ancestral - L1_Ancestral.Insulator,
    L2 + L2.Insulator - L1_Recent - L1_Recent.Insulator,
    LTR + LTR.Insulator - L1_Recent - L1_Recent.Insulator,
    DNA + DNA.Insulator - L1_Recent - L1_Recent.Insulator,
    LTR + LTR.Insulator - L2 - L2.Insulator,
    DNA + DNA.Insulator - L2 - L2.Insulator,
    DNA + DNA.Insulator - LTR - LTR.Insulator,
    levels = modMat)
)
for (i in 2:length(contMat)) colnames(contMat[[i]]) <- colnames(contMat[[1]])
```

# Confidence Intervals

$1 - \alpha$ Confidence intervals were then formed using the `sandwich` estimator for standard errors, setting $\alpha = 0.05 / 6$

```{r, confintCond}
confintCond <- names(contMat) %>%
  lapply(function(x){
    fig3Lm %>%
      update(.~. -1) %>%
      glht(vcov = sandwich, linfct = t(contMat[[x]])) %>%
      confint(level = 1 - 0.05/6) %>%
      extract2("confint") %>%
      as.data.frame() %>%
      add_rownames(var = "Comparison") %>%
      mutate(RE = x)
  }) %>%
  bind_rows %>%
  mutate(Sig = sign(lwr) == sign(upr),
         RE = factor(RE, levels = unique(RE)),
         Comparison = factor(Comparison, levels = unique(Comparison))) 
```

```{r, fig3b, echo=FALSE, fig.width=9}
fig3b <- confintCond %>%
  mutate(Comparison = gsub("_", " ", Comparison),
         Comparison = factor(Comparison, levels = rev(unique(Comparison))),
         RE = gsub(" ", "\n", RE),
         RE = factor(RE, levels = unique(RE))) %>%
  ggplot(aes(x = Estimate, y = Comparison)) +
  geom_point(aes(colour = Sig), size = 0.5) + 
  geom_errorbarh(aes(xmin = lwr, xmax = upr, colour = Sig), 
                 height = 0.5, size = 0.4) +
  geom_vline(xintercept = 0, 
             linetype = 2,
             colour = "grey50", size = 0.2) +
  facet_wrap(~RE, ncol = 6) +
  scale_colour_manual(values = c("black", "red")) +
  guides(colour = FALSE) +
  theme_bw() +
  labs(x = "Difference in logit(p)",
       y = c()) +
  theme(panel.grid = element_blank(),
        panel.margin = unit(0.004,"npc"),
        axis.text = element_text(size = fs),
        axis.title = element_text(size = fs),
        strip.text = element_text(size = fs -1 ))
fig3b
```

# Export of the complete image
```{r, export, echo=FALSE, include=FALSE}
png("Figure3.png", 14.5, 14.5, units = "cm", res = 300)
grid.newpage()
vp1 <- viewport(x=0.085, y=0.55, w=0.915, h=0.48, just=c("left","bottom"))
vp2 <- viewport(x= 0, y=0, w=1, h=0.57, just=c("left","bottom"))
print(fig3b +
        theme(axis.text.y = element_text(size = fs- 2 ), 
              axis.text.x = element_text(size = fs - 1),
              axis.title = element_text(size = fs),
              panel.margin = unit(0.004, "npc")),
      vp=vp2)
print(fig3a +
        theme(axis.title.y = element_text(size = fs - 1),
              legend.text = element_text(size = fs - 2),
              legend.title = element_text(size = fs - 1),
              legend.key.size = unit(4, "mm")), 
      vp=vp1)
grid.text("A", x = 0.02, y = 0.96, default.units = "npc", gp = gpar(fontsize = fs + 4))
grid.text("B", x = 0.02, y = 0.54, default.units = "npc", gp = gpar(fontsize = fs + 4))
dev.off()
```

