---
title: "Figure 4"
author: "Steve Pederson"
date: "26 February 2016"
output: html_document
---
# Introduction
This contains code for the generation of Figure 4, S5 and S6

# Setup

```{r, loadPackages, message=FALSE, warning=FALSE}
library(ggplot2)
library(stringr)
library(reshape2)
library(matrixStats)
library(VennDiagram)
library(knitr)
library(xtable)
library(corrplot)
library(scales)
library(pscl)
library(boot)
library(lmtest)
library(dplyr)
library(magrittr)
library(parallel)
library(funsForLu)
library(spShortcuts)
library(multcomp)
```

Load the data
```{r, loadData}
tpmFile <- file.path("data", "Dec_log2_gene_expression_dataset.txt")
tpmData <- read.delim(tpmFile, sep="\t", stringsAsFactors = FALSE) %>% 
  dplyr::select(ID, length, contains("TPM"), contains("IDs")) %>%
  tbl_df
```

Collect the elements by type and region
```{r, elements}
elements <- grep("IDs", colnames(tpmData), value = TRUE) %>% 
  str_split_fixed(pattern="_", 3) %>%
  as.data.frame(stringsAsFactors = FALSE) %>%
  dplyr::select(type = V1, region = V2) %>%
  lapply(unique) %>%
  lapply(sort)
nRegions <- length(elements$region)
nElements <- length(elements$type)
```

Collect the tissues
```{r, tissues}
tissues <- colnames(tpmData) %>%
  str_subset("TPM") %>%
  str_replace("_TPM", "")
nTissues <- length(tissues)
```

Add columns of logical values to the main data object which denote the element type and region
```{r, addElements}
tpmData %<>%
  bind_cols(sapply(unlist(elements) %>% 
                     as.vector, 
                   function(x){dplyr::select(tpmData, contains(x)) %>% 
                                 rowSums %>% 
                                 as.logical},
                   simplify = FALSE)) %>%
  mutate(nElTypes = alu + L1 + L2 + LTR + mir)
```

# Non-expressed genes
Prior to running the boostrap, a logistic regression model was fit to the data based on the binary variable which considered if a genes was expressed (i.e. TPM > 0) or not (i.e. TPM = 0).
This model was defined as:

$y_{ij} = \alpha_{i} + \beta + \gamma_{j}$
where 

- $\alpha_i$ for $i$={1, 2, ..., 6} is the Tissue specific intercept term
- $\beta$ is the effect on the probability of expression due to gene length
- $\gamma_j$ for $j$ = {1, 2, ..., 20} represents any effects on the probability of expression due each of the 20 combinations of TE and genomic region
- $y_{ij}$ = logit($\pi_{ij}$), where $\pi_{ij}$ is the probability of expression for a gene within Tissue $i$ with TEs defined by $j$

No interaction terms were included in the final model.

```{r, binaryData}
binaryData <- tpmData %>%
  mutate(logLength = log(length)) %>%
  melt(id.vars = c("ID", str_subset(colnames(tpmData), "IDs"), "logLength"),
       measure.vars = str_subset(colnames(tpmData), "TPM"),
       variable.name = "Tissue",
       value.name = "TPM") %>% 
  mutate(Tissue = str_replace(Tissue, "_TPM", ""),
         Tissue = str_replace(Tissue, "_", " "),
         Tissue = str_to_title(Tissue),
         Tissue = as.factor(Tissue),
         TPM = is.finite(TPM)) %>%
  as_data_frame()
```


```{r, fm}
fm <- paste("TPM ~ 0 + Tissue + logLength + ",
            str_subset(colnames(binaryData), "IDs") %>%
              sort %>%
              paste(collapse = " + ")) %>%
  as.formula()
```

```{r, expGlm}
expGlm <- glm(fm, "binomial", binaryData) 
expSum <- expGlm %>%
  glht %>%
  summary(test = adjusted("bonferroni")) %>%
  extract2("test") %>%
  extract(c("coefficients", "sigma", "tstat", "pvalues")) %>%
  set_names(c("Estimate", "Std.Error", "T", "P")) %>%
  as.data.frame() %>%
  add_rownames("Coefficient") 
```

The results from the `glm` are summarised below, with all stated $p$-values being adjusted using Bonferroni's method.

```{r, glmRes, echo=FALSE, include=TRUE}
expSum %>%
  mutate(Sig = aster(P),
         Symbol = c(paste0("$\\alpha_", 1:6, "$"),
                    "$\\beta$",
                    paste0("$\\gamma_{", 1:20, "}$")),
         Coefficient = gsub("Tissue(.+)", "\\1 Tissue", Coefficient)) %>% 
  dplyr::select(Symbol, Coefficient, Estimate, Std.Error, T, P, Sig) %>%
  mutate(P = format(P, digits = 3, scientific = TRUE)) %>%
  kable(digits = 4,
        col.names = c("Model Coefficient", "Term", "Estimate", "SE", "T", "P", " "))
```

```{r, fig4a, echo=FALSE, include=TRUE, fig.width=9, fig.cap = 'T statistics from the logistic regression model plotted as a heatmap for each TE in each Genomic Region'}
fig4a <- expSum %>%
  mutate(Symbol = c(paste0("$\\alpha_", 1:6, "$"),
                    "$\\beta$",
                    paste0("$\\gamma_{", 1:20, "}$")),
         Coefficient = gsub("Tissue(.+)", "\\1 Tissue", Coefficient)) %>% 
  dplyr::select(Symbol, Coefficient, Estimate, T, P) %>%
  filter(grepl("gamma", Symbol),
         P < 0.05) %>%
  mutate(TE = gsub("(.+)_.+_IDs", "\\1", Coefficient),
         TE = capwords(TE),
         TE = gsub("Ltr", "LTR", TE),
         TE = factor(TE, rev(unique(TE))),
         Region = gsub(".+_(.+)_IDs", "\\1", Coefficient),
         Region = gsub("prox", "Proximal\nPromoter", Region),
         Region = gsub("utr(5|3)", "\\1'UTR", Region),
         Region = gsub("cds", "CDS", Region)) %>%
  ggplot(aes(x = Region, y = TE, fill = T)) +
  geom_raster() +
  scale_fill_gradient2(low = rgb(.25, .25, .7),
                       mid = rgb(1, 1, 1),
                       high = rgb(.7, .15, .15)) +
  scale_y_discrete(expand = c(0, 0)) +
  theme_bw() + 
  labs(x = c(),
       y = c(),
       fill = "T\nstatistic") +
  facet_wrap(~Region, scales = "free_x", nrow = 1) +
  theme(axis.ticks.x = element_blank(),
        axis.text.x = element_blank(),
        legend.title.align = 0.5,
        panel.grid = element_blank(),
        panel.margin = unit(0.005, "npc")) 
fig4a
```



# Setup For The Bootstrap Procedure

During the weighted bootstrap procedure, genes will be assigned to groups based on their TE structure.
In order to address any effects on TPM counts due to gene length, the TE groups will be further divided based on length quartiles.
The mean difference in length will also be recorded to ensure that the length distributions are matched.

Additionally, a number of genes contained all five elements, and as this set of `r tpmData %>% filter(alu, L1, L2, LTR, mir) %>% nrow()` genes was clearly higher than would be expected by random, this set with all 5 elements were omitted from the analysis.

```{r, makeBins}
nBins <- 10
tpmData %<>% 
  filter(nElTypes != nElements) %>%
  mutate(lengthBin = findInterval(length, 
                                  quantile(length, 
                                           probs = seq(0, 1, length.out = nBins + 1)), 
                                  rightmost.closed=TRUE),
         logLength = log(length))
```

Confidence intervals will be formed for the presence of each TE within each Genomic Region.
However, in order to control the Type 1 Error at the level $\alpha = 0.05$, the entire set of $m$ intervals needs to be adjusted, with the Boferroni's method being the simplest approach.
Thus a set of $1-\alpha / m$ Confidence intervals will be formed for each genomic region.

```{r}
m <- nTissues*nElements*nRegions
alpha <- 0.05/m
nBoot <- 10000L
testVals <- c("logLength", tissues)
```


An additional function was written to extract the CI from each set of bootstrap samples
```{r}
getCI <- function(x, alpha, combined = FALSE){
  if (!combined) {
    out <- x %>%
      lapply(extract2, "samples") %>%
      lapply(quantile, probs = c(alpha/2, 0.5, 1 - alpha/2)) %>%
      lapply(t) %>%
      lapply(as.data.frame) %>%
      bind_rows() %>%
      set_names(c("lower", "median", "upper")) %>%
      mutate(Tissue = names(x))
  } 
  else {
    out <- x %>%
      lapply(extract2, "samples") %>%
      unlist %>%
      quantile(probs = c(alpha/2, 0.5, 1 - alpha/2)) %>%
      as.list %>%
      as.data.frame %>%
      set_names(c("lower", "median", "upper"))
  }
  out
}
```


In order to provide the number of samples to accurately construct these confidence intervals, the analysis was performed using `r nBoot` bootstrap iterations.

## 3' UTR

### Alu Elements

```{r, aluUtr3Data}
aluUtr3Data <- tpmData %>%
  mutate(
    elBins = dplyr::select(
      tpmData, 
      one_of(grep("alu", elements$type, invert = TRUE, value = TRUE))
    ) %>%
    binsFromCols, 
    elBins = paste(elBins, lengthBin, sep = "_")) %>%
  filter(alu_cds_IDs == 0, 
         alu_prox_IDs == 0, 
         alu_utr5_IDs == 0)
```

```{r, bootAluUtr3, cache = TRUE}
bootAluUtr3 <- mclapply(testVals, 
                        bootMeans, 
                        data = aluUtr3Data, 
                        testIds = filter(aluUtr3Data, alu)$ID, 
                        refIds = filter(aluUtr3Data, !alu)$ID,
                        binCol = "elBins",
                        nBoot = nBoot, 
                        nGenes = 1000L,
                        maxP = 0.01)
names(bootAluUtr3) <- testVals
```

### L1 Elements

```{r, L1Utr3Data}
L1Utr3Data <- tpmData %>%
  mutate(
    elBins = dplyr::select(
      tpmData, 
      one_of(grep("L1", elements$type, invert = TRUE, value = TRUE))
    ) %>%
    binsFromCols, 
    elBins = paste(elBins, lengthBin, sep = "_")) %>%
  filter(L1_cds_IDs == 0, 
         L1_prox_IDs == 0, 
         L1_utr5_IDs == 0)
```

```{r, bootL1Utr3, cache = TRUE}
bootL1Utr3 <- mclapply(testVals, 
                        bootMeans, 
                        data = L1Utr3Data, 
                        testIds = filter(L1Utr3Data, L1)$ID, 
                        refIds = filter(L1Utr3Data, !L1)$ID,
                        binCol = "elBins",
                        nBoot = nBoot, 
                        nGenes = 1000L,
                        maxP = 0.01)
names(bootL1Utr3) <- testVals
```

### L2 Elements

```{r, L2Utr3Data}
L2Utr3Data <- tpmData %>%
  mutate(
    elBins = dplyr::select(
      tpmData, 
      one_of(grep("L2", elements$type, invert = TRUE, value = TRUE))
    ) %>%
    binsFromCols, 
    elBins = paste(elBins, lengthBin, sep = "_")) %>%
  filter(L2_cds_IDs == 0, 
         L2_prox_IDs == 0, 
         L2_utr5_IDs == 0)
```

```{r, bootL2Utr3, cache = TRUE}
bootL2Utr3 <- mclapply(testVals, 
                        bootMeans, 
                        data = L2Utr3Data, 
                        testIds = filter(L2Utr3Data, L2)$ID, 
                        refIds = filter(L2Utr3Data, !L2)$ID,
                        binCol = "elBins",
                        nBoot = nBoot, 
                        nGenes = 1000L,
                        maxP = 0.01)
names(bootL2Utr3) <- testVals
```

### LTR Elements

```{r, LTRUtr3Data}
LTRUtr3Data <- tpmData %>%
  mutate(
    elBins = dplyr::select(
      tpmData, 
      one_of(grep("LTR", elements$type, invert = TRUE, value = TRUE))
    ) %>%
    binsFromCols, 
    elBins = paste(elBins, lengthBin, sep = "_")) %>%
  filter(LTR_cds_IDs == 0, 
         LTR_prox_IDs == 0, 
         LTR_utr5_IDs == 0)
```

```{r, bootLTRUtr3, cache = TRUE}
bootLTRUtr3 <- mclapply(testVals, 
                        bootMeans, 
                        data = LTRUtr3Data, 
                        testIds = filter(LTRUtr3Data, LTR)$ID, 
                        refIds = filter(LTRUtr3Data, !LTR)$ID,
                        binCol = "elBins",
                        nBoot = nBoot, 
                        nGenes = 1000L,
                        maxP = 0.01)
names(bootLTRUtr3) <- testVals
```

### mir Elements

```{r, mirUtr3Data}
mirUtr3Data <- tpmData %>%
  mutate(
    elBins = dplyr::select(
      tpmData, 
      one_of(grep("mir", elements$type, invert = TRUE, value = TRUE))
    ) %>%
    binsFromCols, 
    elBins = paste(elBins, lengthBin, sep = "_")) %>%
  filter(mir_cds_IDs == 0, 
         mir_prox_IDs == 0, 
         mir_utr5_IDs == 0)
```

```{r, bootMirUtr3, cache = TRUE}
bootMirUtr3 <- mclapply(testVals, 
                        bootMeans, 
                        data = mirUtr3Data, 
                        testIds = filter(mirUtr3Data, mir)$ID, 
                        refIds = filter(mirUtr3Data, !mir)$ID,
                        binCol = "elBins",
                        nBoot = nBoot, 
                        nGenes = 1000L,
                        maxP = 0.01)
names(bootMirUtr3) <- testVals
```

## 5' UTR

### Alu Elements

```{r, aluUtr5Data}
aluUtr5Data <- tpmData %>%
  mutate(
    elBins = dplyr::select(
      tpmData, 
      one_of(grep("alu", elements$type, invert = TRUE, value = TRUE))
    ) %>%
    binsFromCols, 
    elBins = paste(elBins, lengthBin, sep = "_")) %>%
  filter(alu_cds_IDs == 0, 
         alu_prox_IDs == 0, 
         alu_utr3_IDs == 0)
```

```{r, bootAluUtr5, cache = TRUE}
bootAluUtr5 <- mclapply(testVals, 
                        bootMeans, 
                        data = aluUtr5Data, 
                        testIds = filter(aluUtr5Data, alu)$ID, 
                        refIds = filter(aluUtr5Data, !alu)$ID,
                        binCol = "elBins",
                        nBoot = nBoot, 
                        nGenes = 1000L,
                        maxP = 0.01)
names(bootAluUtr5) <- testVals
```

### L1 Elements

```{r, L1Utr5Data}
L1Utr5Data <- tpmData %>%
  mutate(
    elBins = dplyr::select(
      tpmData, 
      one_of(grep("L1", elements$type, invert = TRUE, value = TRUE))
    ) %>%
    binsFromCols, 
    elBins = paste(elBins, lengthBin, sep = "_")) %>%
  filter(L1_cds_IDs == 0, 
         L1_prox_IDs == 0, 
         L1_utr3_IDs == 0)
```

```{r, bootL1Utr5, cache = TRUE}
bootL1Utr5 <- mclapply(testVals, 
                        bootMeans, 
                        data = L1Utr5Data, 
                        testIds = filter(L1Utr5Data, L1)$ID, 
                        refIds = filter(L1Utr5Data, !L1)$ID,
                        binCol = "elBins",
                        nBoot = nBoot, 
                        nGenes = 1000L,
                        maxP = 0.01)
names(bootL1Utr5) <- testVals
```

### L2 Elements

```{r, L2Utr5Data}
L2Utr5Data <- tpmData %>%
  mutate(
    elBins = dplyr::select(
      tpmData, 
      one_of(grep("L2", elements$type, invert = TRUE, value = TRUE))
    ) %>%
    binsFromCols, 
    elBins = paste(elBins, lengthBin, sep = "_")) %>%
  filter(L2_cds_IDs == 0, 
         L2_prox_IDs == 0, 
         L2_utr3_IDs == 0)
```

```{r, bootL2Utr5, cache = TRUE}
bootL2Utr5 <- mclapply(testVals, 
                        bootMeans, 
                        data = L2Utr5Data, 
                        testIds = filter(L2Utr5Data, L2)$ID, 
                        refIds = filter(L2Utr5Data, !L2)$ID,
                        binCol = "elBins",
                        nBoot = nBoot, 
                        nGenes = 1000L,
                        maxP = 0.01)
names(bootL2Utr5) <- testVals
```

### LTR Elements

```{r, LTRUtr5Data}
LTRUtr5Data <- tpmData %>%
  mutate(
    elBins = dplyr::select(
      tpmData, 
      one_of(grep("LTR", elements$type, invert = TRUE, value = TRUE))
    ) %>%
    binsFromCols, 
    elBins = paste(elBins, lengthBin, sep = "_")) %>%
  filter(LTR_cds_IDs == 0, 
         LTR_prox_IDs == 0, 
         LTR_utr3_IDs == 0)
```

```{r, bootLTRUtr5, cache = TRUE}
bootLTRUtr5 <- mclapply(testVals, 
                        bootMeans, 
                        data = LTRUtr5Data, 
                        testIds = filter(LTRUtr5Data, LTR)$ID, 
                        refIds = filter(LTRUtr5Data, !LTR)$ID,
                        binCol = "elBins",
                        nBoot = nBoot, 
                        nGenes = 1000L,
                        maxP = 0.01)
names(bootLTRUtr5) <- testVals
```

### mir Elements

```{r, mirUtr5Data}
mirUtr5Data <- tpmData %>%
  mutate(
    elBins = dplyr::select(
      tpmData, 
      one_of(grep("mir", elements$type, invert = TRUE, value = TRUE))
    ) %>%
    binsFromCols, 
    elBins = paste(elBins, lengthBin, sep = "_")) %>%
  filter(mir_cds_IDs == 0, 
         mir_prox_IDs == 0, 
         mir_utr3_IDs == 0)
```

```{r, bootMirUtr5, cache = TRUE}
bootMirUtr5 <- mclapply(testVals, 
                        bootMeans, 
                        data = mirUtr5Data, 
                        testIds = filter(mirUtr5Data, mir)$ID, 
                        refIds = filter(mirUtr5Data, !mir)$ID,
                        binCol = "elBins",
                        nBoot = nBoot, 
                        nGenes = 1000L,
                        maxP = 0.01)
names(bootMirUtr5) <- testVals
```

## Proximal Promoter

### Alu Elements

```{r, aluProxData}
aluProxData <- tpmData %>%
  mutate(
    elBins = dplyr::select(
      tpmData, 
      one_of(grep("alu", elements$type, invert = TRUE, value = TRUE))
    ) %>%
    binsFromCols, 
    elBins = paste(elBins, lengthBin, sep = "_")) %>%
  filter(alu_cds_IDs == 0, 
         alu_utr3_IDs == 0, 
         alu_utr5_IDs == 0)
```

```{r, bootAluProx, cache = TRUE}
bootAluProx <- mclapply(testVals, 
                        bootMeans, 
                        data = aluProxData, 
                        testIds = filter(aluProxData, alu)$ID, 
                        refIds = filter(aluProxData, !alu)$ID,
                        binCol = "elBins",
                        nBoot = nBoot, 
                        nGenes = 1000L,
                        maxP = 0.01)
names(bootAluProx) <- testVals
```

### L1 Elements

```{r, L1ProxData}
L1ProxData <- tpmData %>%
  mutate(
    elBins = dplyr::select(
      tpmData, 
      one_of(grep("L1", elements$type, invert = TRUE, value = TRUE))
    ) %>%
    binsFromCols, 
    elBins = paste(elBins, lengthBin, sep = "_")) %>%
  filter(L1_cds_IDs == 0, 
         L1_utr3_IDs == 0, 
         L1_utr5_IDs == 0)
```

```{r, bootL1Prox, cache = TRUE}
bootL1Prox <- mclapply(testVals, 
                        bootMeans, 
                        data = L1ProxData, 
                        testIds = filter(L1ProxData, L1)$ID, 
                        refIds = filter(L1ProxData, !L1)$ID,
                        binCol = "elBins",
                        nBoot = nBoot, 
                        nGenes = 1000L,
                        maxP = 0.01)
names(bootL1Prox) <- testVals
```

### L2 Elements

```{r, L2ProxData}
L2ProxData <- tpmData %>%
  mutate(
    elBins = dplyr::select(
      tpmData, 
      one_of(grep("L2", elements$type, invert = TRUE, value = TRUE))
    ) %>%
    binsFromCols, 
    elBins = paste(elBins, lengthBin, sep = "_")) %>%
  filter(L2_cds_IDs == 0, 
         L2_utr3_IDs == 0, 
         L2_utr5_IDs == 0)
```

```{r, bootL2Prox, cache = TRUE}
bootL2Prox <- mclapply(testVals, 
                        bootMeans, 
                        data = L2ProxData, 
                        testIds = filter(L2ProxData, L2)$ID, 
                        refIds = filter(L2ProxData, !L2)$ID,
                        binCol = "elBins",
                        nBoot = nBoot, 
                        nGenes = 1000L,
                        maxP = 0.01)
names(bootL2Prox) <- testVals
```

### LTR Elements

```{r, LTRProxData}
LTRProxData <- tpmData %>%
  mutate(
    elBins = dplyr::select(
      tpmData, 
      one_of(grep("LTR", elements$type, invert = TRUE, value = TRUE))
    ) %>%
    binsFromCols, 
    elBins = paste(elBins, lengthBin, sep = "_")) %>%
  filter(LTR_cds_IDs == 0, 
         LTR_utr3_IDs == 0, 
         LTR_utr5_IDs == 0)
```

```{r, bootLTRProx, cache = TRUE}
bootLTRProx <- mclapply(testVals, 
                        bootMeans, 
                        data = LTRProxData, 
                        testIds = filter(LTRProxData, LTR)$ID, 
                        refIds = filter(LTRProxData, !LTR)$ID,
                        binCol = "elBins",
                        nBoot = nBoot, 
                        nGenes = 1000L,
                        maxP = 0.01)
names(bootLTRProx) <- testVals
```

### mir Elements

```{r, mirProxData}
mirProxData <- tpmData %>%
  mutate(
    elBins = dplyr::select(
      tpmData, 
      one_of(grep("mir", elements$type, invert = TRUE, value = TRUE))
    ) %>%
    binsFromCols, 
    elBins = paste(elBins, lengthBin, sep = "_")) %>%
  filter(mir_cds_IDs == 0, 
         mir_utr3_IDs == 0, 
         mir_utr5_IDs == 0)
```

```{r, bootMirProx, cache = TRUE}
bootMirProx <- mclapply(testVals, 
                        bootMeans, 
                        data = mirProxData, 
                        testIds = filter(mirProxData, mir)$ID, 
                        refIds = filter(mirProxData, !mir)$ID,
                        binCol = "elBins",
                        nBoot = nBoot, 
                        nGenes = 1000L,
                        maxP = 0.01)
names(bootMirProx) <- testVals
```


## CDS regions

### Alu Elements

```{r, aluCDSData}
aluCDSData <- tpmData %>%
  mutate(
    elBins = dplyr::select(
      tpmData, 
      one_of(grep("alu", elements$type, invert = TRUE, value = TRUE))
    ) %>%
    binsFromCols, 
    elBins = paste(elBins, lengthBin, sep = "_")) %>%
  filter(alu_prox_IDs == 0, 
         alu_utr3_IDs == 0, 
         alu_utr5_IDs == 0)
```

```{r, bootAluCDS, cache = TRUE}
bootAluCDS <- mclapply(testVals, 
                        bootMeans, 
                        data = aluCDSData, 
                        testIds = filter(aluCDSData, alu)$ID, 
                        refIds = filter(aluCDSData, !alu)$ID,
                        binCol = "elBins",
                        nBoot = nBoot, 
                        nGenes = 45L,
                        maxP = 0.1)
names(bootAluCDS) <- testVals
```

### L1 Elements

```{r, L1CDSData}
L1CDSData <- tpmData %>%
  mutate(
    elBins = dplyr::select(
      tpmData, 
      one_of(grep("L1", elements$type, invert = TRUE, value = TRUE))
    ) %>%
    binsFromCols, 
    elBins = paste(elBins, lengthBin, sep = "_")) %>%
  filter(L1_prox_IDs == 0, 
         L1_utr3_IDs == 0, 
         L1_utr5_IDs == 0)
```

```{r, bootL1CDS, cache = TRUE}
bootL1CDS <- mclapply(testVals, 
                        bootMeans, 
                        data = L1CDSData, 
                        testIds = filter(L1CDSData, L1)$ID, 
                        refIds = filter(L1CDSData, !L1)$ID,
                        binCol = "elBins",
                        nBoot = nBoot, 
                        nGenes = 45L,
                        maxP = 0.1)
names(bootL1CDS) <- testVals
```

### L2 Elements

```{r, L2CDSData}
L2CDSData <- tpmData %>%
  mutate(
    elBins = dplyr::select(
      tpmData, 
      one_of(grep("L2", elements$type, invert = TRUE, value = TRUE))
    ) %>%
    binsFromCols, 
    elBins = paste(elBins, lengthBin, sep = "_")) %>%
  filter(L2_prox_IDs == 0, 
         L2_utr3_IDs == 0, 
         L2_utr5_IDs == 0)
```

```{r, bootL2CDS, cache = TRUE}
bootL2CDS <- mclapply(testVals, 
                        bootMeans, 
                        data = L2CDSData, 
                        testIds = filter(L2CDSData, L2)$ID, 
                        refIds = filter(L2CDSData, !L2)$ID,
                        binCol = "elBins",
                        nBoot = nBoot, 
                        nGenes = 55L,
                        maxP = 0.1)
names(bootL2CDS) <- testVals
```

### LTR Elements

```{r, LTRCDSData}
LTRCDSData <- tpmData %>%
  mutate(
    elBins = dplyr::select(
      tpmData, 
      one_of(grep("LTR", elements$type, invert = TRUE, value = TRUE))
    ) %>%
    binsFromCols, 
    elBins = paste(elBins, lengthBin, sep = "_")) %>%
  filter(LTR_prox_IDs == 0, 
         LTR_utr3_IDs == 0, 
         LTR_utr5_IDs == 0)
```

```{r, bootLTRCDS, cache = TRUE}
bootLTRCDS <- mclapply(testVals, 
                        bootMeans, 
                        data = LTRCDSData, 
                        testIds = filter(LTRCDSData, LTR)$ID, 
                        refIds = filter(LTRCDSData, !LTR)$ID,
                        binCol = "elBins",
                        nBoot = nBoot, 
                        nGenes = 33L,
                        maxP = 0.01)
names(bootLTRCDS) <- testVals
```

### mir Elements

```{r, mirCDSData}
mirCDSData <- tpmData %>%
  mutate(
    elBins = dplyr::select(
      tpmData, 
      one_of(grep("mir", elements$type, invert = TRUE, value = TRUE))
    ) %>%
    binsFromCols, 
    elBins = paste(elBins, lengthBin, sep = "_")) %>%
  filter(mir_prox_IDs == 0, 
         mir_utr3_IDs == 0, 
         mir_utr5_IDs == 0)
```

```{r, bootMirCDS, cache = TRUE}
bootMirCDS <- mclapply(testVals, 
                        bootMeans, 
                        data = mirCDSData, 
                        testIds = filter(mirCDSData, mir)$ID, 
                        refIds = filter(mirCDSData, !mir)$ID,
                        binCol = "elBins",
                        nBoot = nBoot, 
                        nGenes = 35L,
                        maxP = 0.1)
names(bootMirCDS) <- testVals
```

# Confidence Intervals for Length Differences

This step is a control to check that the length distributions were suitably paired.
Any intervals that don't contain zero will indicate that the two sets of genes will have different length distributions and the comparison will need to be reanalysed

```{r}
confIntLen <- list(bootAluUtr3, bootAluUtr5, bootAluProx, bootAluCDS,
                   bootL1Utr3, bootL1Utr5, bootL1Prox, bootL1CDS,
                   bootL2Utr3, bootL2Utr5, bootL2Prox, bootL2CDS,
                   bootLTRUtr3, bootLTRUtr5, bootLTRProx, bootLTRCDS,
                   bootMirUtr3, bootMirUtr5, bootMirProx, bootMirCDS) %>%
  lapply(extract, "logLength") %>%
  lapply(getCI, alpha = 0.05, combined = TRUE) %>%
  bind_rows %>%
  mutate(Region = rep(c("3'UTR", "5'UTR","Proximal Promoter", "CDS"), 
                      times = nElements),
         TE = rep(elements$type, each = nRegions))
```

```{r, figS5, echo=FALSE, include=TRUE, fig.width=8}
figS5 <- confIntLen %>%
  mutate(TE = capwords(TE),
         TE = factor(TE, rev(unique(TE)))) %>%
  ggplot(aes(x = median, y = TE)) +
  geom_errorbarh(aes(xmin = lower, xmax = upper), height = 0.5) +
  facet_wrap(~Region, nrow=1, scales = "free_x") +
  geom_vline(xintercept = 0, colour = "grey50", linetype = 2) +
  theme_bw() +
  labs(x = "95% CI for the diference in log(Length)") 
figS5
```

As type II errors are the primary concern here, no adjustment was applied for multiple comparisons.
Zero didn't lie outside of any of these intervals, and as such the subsequent results for the bootstrap can be deemed to be valid.

```{r, exportFigS5, echo = FALSE, include=TRUE, message=FALSE}
png("FigureS5.png", height = 10, width = 14.5, units = "cm", res = 300)
fs <- 9
figS5 +
  theme(axis.text = element_text(size = fs - 0.5),
        axis.title = element_text(size = fs + 1),
        strip.text = element_text(size = fs + 1),
        panel.margin = unit(1, "mm"))
dev.off()
```


# Confidence Intervals for Elements & Regions

## Combined Across Tissues

```{r, confIntCombined}
m <- nRegions*nElements
confIntCombined <- list(bootAluUtr3, bootAluUtr5, bootAluProx, bootAluCDS,
                        bootL1Utr3, bootL1Utr5, bootL1Prox, bootL1CDS,
                        bootL2Utr3, bootL2Utr5, bootL2Prox, bootL2CDS,
                        bootLTRUtr3, bootLTRUtr5, bootLTRProx, bootLTRCDS,
                        bootMirUtr3, bootMirUtr5, bootMirProx, bootMirCDS) %>%
  lapply(extract, tissues) %>%
  lapply(getCI, 
         alpha = 0.05/m, 
         combined = TRUE) %>%
  bind_rows() %>%
  mutate(TE = rep(elements$type, each = nRegions),
         Region = rep(c("3'UTR", "5'UTR", "Proximal Promoter", "CDS"),
                      times = nElements),
         Sig = sign(lower) == sign(upper))
```

```{r}
confIntCombined %>%
  mutate(TE = factor(TE, rev(unique(TE)))) %>%
  ggplot(aes(x = median, y = TE, colour = Sig)) +
  geom_errorbarh(aes(xmin = lower, xmax =upper)) +
  geom_vline(xintercept = 0) +
  theme_bw() + 
  scale_color_manual(values = c("black", "red")) +
  facet_wrap(~Region, nrow = 1, scales = "free_x") +
  guides(colour = FALSE) +
  labs(x = "FWER-adjusted 95% Confidence Intervals",
       y = c(),
       title = "Bootstrap Confidence Intervals For log(TPM)")
```

No sigificant effects were detected across all tissues, although the restrictive power of a Bonferroni correction may be responsible for this, especially with Alu elements.

## Tissue Specific Effects

As there appeared to be a bias towards upregulatory effects due to the presence of an Alu element, tissue-specific effects were investigated for all elements

```{r, confIntTissues}
m <- nRegions*nElements*nTissues
confIntTissues <- list(bootAluUtr3, bootAluUtr5, bootAluProx, bootAluCDS,
                        bootL1Utr3, bootL1Utr5, bootL1Prox, bootL1CDS,
                        bootL2Utr3, bootL2Utr5, bootL2Prox, bootL2CDS,
                        bootLTRUtr3, bootLTRUtr5, bootLTRProx, bootLTRCDS,
                        bootMirUtr3, bootMirUtr5, bootMirProx, bootMirCDS) %>%
  lapply(extract, tissues) %>%
  lapply(getCI, 
         alpha = 0.05/m) %>%
  bind_rows() %>%
  mutate(TE = rep(elements$type, 
                  each = nRegions*nTissues),
         Region = c("3'UTR", "5'UTR", "Proximal Promoter", "CDS") %>%
           rep(times =nElements) %>%
           rep(each = nTissues),
         Sig = sign(lower) == sign(upper))
```

```{r, plotConfIntTissues, echo = FALSE, include=TRUE, fig.height=7, fig.width=10}
fig4b <- confIntTissues %>%
  mutate(Tissue = capwords(Tissue, sep = "_"),
         Tissue = factor(Tissue, rev(unique(sort(Tissue)))),
         TE = capwords(TE),
         TE = gsub("Ltr", "LTR", TE),
         Region = gsub(" ", "\n", Region),
         lower = lower/log(2),
         median = median / log(2),
         upper = upper / log(2)) %>%
  ggplot(aes(x = median, y = Tissue, colour = Sig)) +
  geom_point(size = 1)+
  geom_errorbarh(aes(xmin = lower,xmax =upper),
                 height = 0.5) +
  scale_color_manual(values = c("black", "red")) +
  geom_vline(xintercept = 0, linetype = 2, colour = "grey70", size = 0.5) +
  theme_bw() + 
  facet_grid(TE~Region, scales = "free_x") +
  guides(colour = FALSE) +
  labs(x = "FWER-adjusted 95% Confidence Intervals",
       y = c(),
       title = "Tissue-Specific Bootstrap Confidence Intervals For log(TPM)")
fig4b
```


```{r, exportPNG, echo=FALSE, include=FALSE}
png("Figure4.png", height = 14.5, width = 14.5, units = "cm", res = 300)
fs <- 9
fig4b +
  labs(title = c(),
       x = expression(paste("Change in ", log[2],"(TPM)"))) +
  theme(axis.text = element_text(size = fs - 0.5),
        axis.title = element_text(size = fs + 1),
        strip.text = element_text(size = fs + 1),
        panel.margin = unit(1, "mm"))
dev.off()
```


```{r, confIntTissuesNoAdj}
confIntTissuesNoAdj <- list(bootAluUtr3, bootAluUtr5, bootAluProx, bootAluCDS,
                            bootL1Utr3, bootL1Utr5, bootL1Prox, bootL1CDS,
                            bootL2Utr3, bootL2Utr5, bootL2Prox, bootL2CDS,
                            bootLTRUtr3, bootLTRUtr5, bootLTRProx, bootLTRCDS,
                            bootMirUtr3, bootMirUtr5, bootMirProx, bootMirCDS) %>%
  lapply(extract, tissues) %>%
  lapply(getCI, 
         alpha = 0.05) %>%
  bind_rows() %>%
  mutate(TE = rep(elements$type, 
                  each = nRegions*nTissues),
         Region = c("3'UTR", "5'UTR", "Proximal Promoter", "CDS") %>%
           rep(times =nElements) %>%
           rep(each = nTissues),
         Sig = sign(lower) == sign(upper))
```

```{r, plotConfIntTissuesNoAdj, echo = FALSE, include=TRUE, fig.height=7, fig.width=10}
figS6 <- confIntTissuesNoAdj %>%
  mutate(Tissue = capwords(Tissue, sep = "_"),
         Tissue = factor(Tissue, rev(unique(sort(Tissue)))),
         TE = capwords(TE),
         TE = gsub("Ltr", "LTR", TE),
         Region = gsub(" ", "\n", Region),
         lower = lower/log(2),
         median = median / log(2),
         upper = upper / log(2)) %>%
  ggplot(aes(x = median, y = Tissue, colour = Sig)) +
  geom_point(size = 1)+
  geom_errorbarh(aes(xmin = lower,xmax =upper),
                 height = 0.5) +
  scale_color_manual(values = c("black", "red")) +
  geom_vline(xintercept = 0, linetype = 2, colour = "grey70", size = 0.5) +
  theme_bw() + 
  facet_grid(TE~Region, scales = "free_x") +
  guides(colour = FALSE) +
  labs(x = "Raw 95% Confidence Intervals",
       y = c(),
       title = "Tissue-Specific Bootstrap Confidence Intervals For log(TPM)")
figS6
```


```{r, exportFigS6, echo=FALSE, include=FALSE}
png("FigureS6.png", height = 12, width = 14.5, units = "cm", res = 300)
fs <- 9
figS6 +
  labs(title = c(),
       x = expression(paste("Change in ", log[2],"(TPM)"))) +
  theme(axis.text = element_text(size = fs - 0.5),
        axis.title = element_text(size = fs + 1),
        strip.text = element_text(size = fs + 1),
        panel.margin = unit(1, "mm"))
dev.off()
```


```{r, exportFig4, echo=FALSE, include=FALSE}
png("Figure4.png", width = 14.5, height = 16, units = "cm", res = 300)
grid.newpage()
vp1 <- viewport(x=-0.02, y=0.75, w=0.973, h=0.25, just=c("left","bottom"))
vp2 <- viewport(x= 0.01, y=0, w=0.99, h=0.75, just=c("left","bottom"))
print(fig4a + 
        theme(legend.title = element_text(size = fs),
              legend.text = element_text(size = fs ),
              legend.key.height = unit(0.025, "npc"),
              legend.position = "left",
              axis.text.y = element_text(size = fs - 0.5)),
      vp=vp1)
print(fig4b + 
        labs(title = c(),
             x = c()) + 
        theme(axis.text = element_text(size = fs),
              panel.margin = unit(0.005, "npc")),
      vp=vp2)
grid.text("A", x = 0.03, y = 0.97, default.units = "npc", gp = gpar(fontsize = fs + 4))
grid.text("B", x = 0.03, y = 0.72, default.units = "npc", gp = gpar(fontsize = fs + 4))
dev.off()
```




```{r}
save.image("Figure4.RData")
```

```{r}
sessionInfo()
```

