---
title: "Distribution of REs and RETEs"
author: "Steve Pederson & Lu Zeng"
date: "17 February 2016"
output: html_document
---

# Load the packages and the data

```{r, warning=FALSE, message=FALSE}
library(ggplot2)
library(MASS)
library(matrixStats)
library(xtable)
library(multcomp)
library(sandwich)
library(boot)
library(dplyr)
library(reshape2)
library(readr)
library(knitr)
library(magrittr)
library(limma)
library(spShortcuts)
library(car)
library(grid)
```

# Introduction

This script:

- Uses proportions transformed to the logit scale
- Tests for normality of data and homoscedasticity (i.e. equal variance across samples)
- Analyses the data using $p(TE\mid RE)$
- Generates Figure 2 & Table S1


```{r, loadData}
fig2 <- file.path("data", "Figure2.csv") %>%
  read_csv(col_names = FALSE) %>%
  set_colnames(c("Region", "RE", "Tissue", "bpRETE", "bpRE")) %>%
  mutate(Region = gsub("([35])'UTR", "UTR\\1", Region),
         Region = gsub("CDS-", "CDS", Region),
         Region = factor(Region, levels=c("UTR5", "CDSexon", "CDSintron",  
                                          "UTR3", "Intergenic")),
         RE = gsub("Repressor", "Repressed_region", RE),
         RE = factor(RE, levels= c("Active_promoter","Weak_promoter", "Strong_enhancer", 
                                   "Weak_enhancer", "Repressed_region", "Insulator")),
         Tissue = as.factor(Tissue),
         bpRETE = bpRETE / 100,
         bpRE = bpRE/100,
         `bpTE|RE` = bpRETE / bpRE,
         logitRETE = logit(bpRETE),
         logitRE = logit(bpRE),
         `logitTE|RE` = logit(`bpTE|RE`))
```

# Raw Data

```{r, fig2a, echo=FALSE, fig.height=6, fig.width=8, fig.cap='The estimated probabilities of a base being a Regulatory Element with each genomic region, and the estimated probabilities of a base within each Regulatory Element being a Transposable Element. Error bars indicate one standard error.'}
fs <- 8
fig2a <- fig2 %>% 
  group_by(Region, RE) %>% 
  summarise(n = n(),
            mnRE = mean(logitRE),
            seRE = sd(logitRE) / sqrt(n),
            `mnTE|RE` = mean(`logitTE|RE`), 
            `seTE|RE` = sd(`logitTE|RE`) / sqrt(n)) %>% 
  ungroup %>%
  mutate(lowerRE = mnRE - seRE,
         `lowerTE|RE` = `mnTE|RE` - `seTE|RE`,
         upperRE = mnRE + seRE,
         `upperTE|RE` = `mnTE|RE` + `seTE|RE`) %>%
  (function(x){
    d1 = melt(x, id.vars = c("Region", "RE"),
              measure.vars = c("mnRE", "mnTE|RE"),
              variable.name = "Class", 
              value.name = "logitPi")
    d2 = melt(x, id.vars = c("Region", "RE"),
              measure.vars = c("lowerRE", "lowerTE|RE"),
              variable.name = "Class", 
              value.name = "logitLower")
    d3 = melt(x, id.vars = c("Region", "RE"),
              measure.vars = c("upperRE", "upperTE|RE"),
              variable.name = "Class", 
              value.name = "logitUpper")
    cbind(d1, logitLower = d2$logitLower, logitUpper = d3$logitUpper)
  }) %>%
  mutate(Region = gsub("UTR(3|5)", "\\1'UTR", Region),
         Region = gsub("CDS", "CDS ", Region),
         Region = factor(Region, levels = unique(Region)),
         RE = gsub("_", " ", RE),
         RE = capwords(RE),
         RE = gsub(" ", "\n", RE),
         RE = factor(RE, levels = unique(RE)),
         Class = gsub("mnRE", "p(RE)", Class),
         Class = gsub("mnTE\\|RE", "p(TE|RE)", Class),
         pi = binomial()$linkinv(logitPi),
         lower = binomial()$linkinv(logitLower),
         upper = binomial()$linkinv(logitUpper)) %>%
  ggplot(aes(x = Region, y = pi, fill = Region)) + 
  geom_bar(stat = "identity") + 
  geom_errorbar(aes(ymin = lower, ymax = upper),
                width = 0.4) +
  facet_grid(Class~RE, scale = "free_y") +
  scale_y_continuous(expand = c(0.02,0)) +
  theme_bw() +
  labs(x = c(), 
       y = "Probability Estimate") +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.y = element_text(size = fs),
        axis.text.y = element_text(size = fs-1),
        strip.text = element_text(size = fs-1),
        legend.text = element_text(size = fs),
        legend.title = element_text(size = fs, 
                                    face = "bold"),
        legend.position = "top",
        legend.key.size = unit(4, "mm"),
        panel.margin = unit(.004, "npc"),
        panel.grid = element_line(colour = "grey50")) 
fig2a
```

# Analysis of conditional probabilities

First an initial model was specified for the conditional probabilities, including interaction terms between each region and each RE.
```{r, lmCond}
lmCond <- lm(`logitTE|RE` ~ Region*RE, fig2)
```

The Shapiro-Wilk Test was used to assess normality, and this assumption was not violated.
```{r}
lmCond %>%
  resid() %>%
  shapiro.test()
```


Levene's Test was performed to assess heteroscedasticity
```{r}
lmCond %>%
  leveneTest(center = "mean")
```
As this assumption has been violated, all standard errors will be calculated using the sandwich estimator.

```{r, resultsLmCond}
resLmCond <- lmCond %>% 
  glht(vcov = sandwich) %>%
  summary(test = adjusted("none")) %>%
  extract2("test") %>%
  extract(c("coefficients", "sigma", "tstat", "pvalues")) %>%
  as.data.frame() %>%
  add_rownames("Coefficient") %>%
  rename(Estimate = coefficients,
         `Std. Error` = sigma,
         T = tstat,
         P = pvalues) %>%
  mutate(Coefficient = gsub("(Region|RE)", "", Coefficient),
         Coefficient = gsub("_", " ", Coefficient),
         Coefficient = capwords(Coefficient),
         Coefficient = gsub("Utr(3|5)", "\\1'UTR", Coefficient),
         Coefficient = gsub("Cds", "CDS-", Coefficient),
         Coefficient = gsub(":(.)", ":\\U\\1", Coefficient, perl = TRUE),
         adjP = p.adjust(P, "bonferroni"),
         ` ` = aster(adjP)) 
```

```{r, kableLmCond, echo=FALSE, include = TRUE}
resLmCond %>%
  mutate(P = format(P, digits = 3, scientific = TRUE),
         adjP = format(adjP, digits = 3, scientific = TRUE)) %>%
  kable(digits = c(0, 3, 4, 3, 0, 0, 0),
        align = c("l", "r")[c(1, 2, 2, 2, 2, 2, 1)],
        caption = paste("Results from analysis of conditional probabilities across all regulatory elements and genomic regions using a linear regression model.",
                        "The values being analysed are the logit transformed proportions of bases, such that a positive value corresponds to an increase in probability.",
                        "Bonferroni's adjustment has been applied to adjusted $p$-values, and the sandwich estimator was utilised for robust standard error estimation.",
                        "Asterisks indicate significance as per the standard conventions of R"))
```

# Confidence Intervals

In order to make comparisons between $p(TE \mid RE)$ estimates by region and element, a matrix of pair-wise comparisons was formed manually.

```{r, contMat}
modMat <- lmCond %>%
  update(.~. -1) %>%
  model.matrix
colnames(modMat) <- gsub("(Region|RE)", "", colnames(modMat))
colnames(modMat) <- gsub(":", ".", colnames(modMat))
contMat <- list(
  `Active Promoter` = makeContrasts(
    CDSexon - UTR5, 
    CDSexon - CDSintron,
    CDSexon - UTR3,
    CDSexon - Intergenic,
    UTR3 - UTR5,
    UTR3 - Intergenic,
    UTR3 - CDSintron,
    Intergenic - UTR5,
    Intergenic - CDSintron,
    CDSintron - UTR5,
    levels = modMat),
  `Weak Promoter` = makeContrasts(
    CDSexon + CDSexon.Weak_promoter - UTR5, 
    CDSexon + CDSexon.Weak_promoter -  CDSintron - CDSintron.Weak_promoter,
    CDSexon + CDSexon.Weak_promoter - UTR3 - UTR3.Weak_promoter,
    CDSexon + CDSexon.Weak_promoter - Intergenic - Intergenic.Weak_promoter,
    UTR3 + UTR3.Weak_promoter - UTR5,
    UTR3 + UTR3.Weak_promoter - Intergenic - Intergenic.Weak_promoter,
    UTR3 + UTR3.Weak_promoter -  CDSintron - CDSintron.Weak_promoter,
    Intergenic + Intergenic.Weak_promoter - UTR5,
    Intergenic + Intergenic.Weak_promoter - CDSintron - CDSintron.Weak_promoter,
    CDSintron + CDSintron.Weak_promoter - UTR5,
    levels = modMat),
  `Strong Enhancer` = makeContrasts(
    CDSexon + CDSexon.Strong_enhancer - UTR5, 
    CDSexon + CDSexon.Strong_enhancer -  CDSintron - CDSintron.Strong_enhancer,
    CDSexon + CDSexon.Strong_enhancer - UTR3 - UTR3.Strong_enhancer,
    CDSexon + CDSexon.Strong_enhancer - Intergenic - Intergenic.Strong_enhancer,
    UTR3 + UTR3.Strong_enhancer - UTR5,
    UTR3 + UTR3.Strong_enhancer - Intergenic - Intergenic.Strong_enhancer,
    UTR3 + UTR3.Strong_enhancer -  CDSintron - CDSintron.Strong_enhancer,
    Intergenic + Intergenic.Strong_enhancer - UTR5,
    Intergenic + Intergenic.Strong_enhancer - CDSintron - CDSintron.Strong_enhancer,
    CDSintron + CDSintron.Strong_enhancer - UTR5,
    levels = modMat),
  `Weak Enhancer` = makeContrasts(
    CDSexon + CDSexon.Weak_enhancer - UTR5, 
    CDSexon + CDSexon.Weak_enhancer -  CDSintron - CDSintron.Weak_enhancer,
    CDSexon + CDSexon.Weak_enhancer - UTR3 - UTR3.Weak_enhancer,
    CDSexon + CDSexon.Weak_enhancer - Intergenic - Intergenic.Weak_enhancer,
    UTR3 + UTR3.Weak_enhancer - UTR5,
    UTR3 + UTR3.Weak_enhancer - Intergenic - Intergenic.Weak_enhancer,
    UTR3 + UTR3.Weak_enhancer -  CDSintron - CDSintron.Weak_enhancer,
    Intergenic + Intergenic.Weak_enhancer - UTR5,
    Intergenic + Intergenic.Weak_enhancer - CDSintron - CDSintron.Weak_enhancer,
    CDSintron + CDSintron.Weak_enhancer - UTR5,
    levels = modMat),
  `Repressed Region` = makeContrasts(
    CDSexon + CDSexon.Repressed_region - UTR5, 
    CDSexon + CDSexon.Repressed_region -  CDSintron - CDSintron.Repressed_region,
    CDSexon + CDSexon.Repressed_region - UTR3 - UTR3.Repressed_region,
    CDSexon + CDSexon.Repressed_region - Intergenic - Intergenic.Repressed_region,
    UTR3 + UTR3.Repressed_region - UTR5,
    UTR3 + UTR3.Repressed_region - Intergenic - Intergenic.Repressed_region,
    UTR3 + UTR3.Repressed_region -  CDSintron - CDSintron.Repressed_region,
    Intergenic + Intergenic.Repressed_region - UTR5,
    Intergenic + Intergenic.Repressed_region - CDSintron - CDSintron.Repressed_region,
    CDSintron + CDSintron.Repressed_region - UTR5,
    levels = modMat),
  `Insulator` = makeContrasts(
    CDSexon + CDSexon.Insulator - UTR5, 
    CDSexon + CDSexon.Insulator -  CDSintron - CDSintron.Insulator,
    CDSexon + CDSexon.Insulator - UTR3 - UTR3.Insulator,
    CDSexon + CDSexon.Insulator - Intergenic - Intergenic.Insulator,
    UTR3 + UTR3.Insulator - UTR5,
    UTR3 + UTR3.Insulator - Intergenic - Intergenic.Insulator,
    UTR3 + UTR3.Insulator -  CDSintron - CDSintron.Insulator,
    Intergenic + Intergenic.Insulator - UTR5,
    Intergenic + Intergenic.Insulator - CDSintron - CDSintron.Insulator,
    CDSintron + CDSintron.Insulator - UTR5,
    levels = modMat)
  
)
for (i in 2:6) colnames(contMat[[i]]) <- colnames(contMat[[1]])
```

As there were 6 regulatory elements under investigation, $1- \alpha$ Confidence Intervals were generated within each RE using the package `multcomp`.
These intervals will control the FWER at a given level within each RE, thus $\alpha = 0.05 / 6$ was set for each RE.

```{r, confintCond}
confintCond <- names(contMat) %>%
  lapply(function(x){
    lmCond %>%
      update(.~. -1) %>%
      glht(vcov = sandwich, linfct = t(contMat[[x]])) %>%
      confint(level = 1 - 0.05/6) %>%
      extract2("confint") %>%
      as.data.frame() %>%
      add_rownames(var = "Comparison") %>%
      mutate(RE = x)
  }) %>%
  bind_rows %>%
  mutate(Sig = sign(lwr) == sign(upr),
         RE = factor(RE, levels = unique(RE)),
         Comparison = factor(Comparison, levels = unique(Comparison)))
```

These Confidence Intervals were then plotted across the set of REs.

```{r, fig2b, echo=FALSE, include=TRUE, fig.height=6, fig.width=9, fig.cap="FWER-adjusted 95\\% Confidence Intervals for the difference in probabilities. Intervals which do not contain zero are coloured red, indicating a rejection of the null hypothesis that there is no difference between these two probabilities."}
fig2b <- confintCond %>%
  mutate(Comparison = gsub("CDS", "CDS ", Comparison),
         Comparison = gsub("UTR(3|5)", "\\1'UTR", Comparison),
         Comparison = factor(Comparison, levels = rev(unique(Comparison))),
         RE = gsub(" ", "\n", RE),
         RE = factor(RE, levels = unique(RE))) %>%
  ggplot(aes(x = Estimate, y = Comparison)) +
  geom_point(aes(colour = Sig), size = 0.5) + 
  geom_errorbarh(aes(xmin = lwr, xmax = upr, colour = Sig), 
                 height = 0.5, size = 0.4) +
  geom_vline(xintercept = 0, 
             linetype = 2,
             colour = "grey50", size = 0.2) +
  facet_wrap(~RE, ncol = 6, scales = "free_x") +
  scale_colour_manual(values = c("black", "red")) +
  guides(colour = FALSE) +
  theme_bw() +
  labs(x = "Difference in logit(p)",
       y = c()) +
  theme(panel.grid = element_blank(),
        panel.margin = unit(0.004,"npc"),
        axis.text = element_text(size = fs),
        axis.title = element_text(size = fs),
        strip.text = element_text(size = fs -1 ))
fig2b
```

## Export the final combined figure
```{r, exportPng, eval=FALSE, echo=FALSE, include=FALSE}
fs <- 8
png("Figure2.png", 14.5, 14.5, units = "cm", res = 300)
grid.newpage()
vp1 <- viewport(x=0.085, y=0.45, w=0.915, h=0.57, just=c("left","bottom"))
vp2 <- viewport(x= 0, y=0, w=0.955, h=0.47, just=c("left","bottom"))
print(fig2a + 
        theme(axis.title.y = element_text(size = fs - 1),
              legend.text = element_text(size = fs - 2),
              legend.title = element_text(size = fs - 1)), 
      vp=vp1)
print(fig2b +
        theme(axis.text.y = element_text(size = fs - 2),
              axis.text.x = element_text(size = fs - 1)), 
      vp=vp2)
grid.text("A", x = 0.03, y = 0.96, default.units = "npc", gp = gpar(fontsize = fs + 4))
grid.text("B", x = 0.03, y = 0.43, default.units = "npc", gp = gpar(fontsize = fs + 4))
dev.off()
```


